---
title: "STAT380_Mini-Project2"
author: "Sammit Bal, Thaddeus Poff, Jason Zhang"
date: "2025-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter
```{r}
library(tidyverse)
library(FNN)
library(readxl)
COD_dataset <- read_excel("./CODGames2_mp.xlsx")
```



## Task 1

**Question:** Explain why you should avoid using the function `na.omit()` to remove observations (rows) with missing data for the `CODGames2_mp` dataset. 


We should avoid using the `na.omit()` function because it will remove an unnecessary amount of observations which still have valuable data in them. In this data set we have a lot of unused variables because of the types of game modes that are played in COD. Certain game modes keep track of more data than others. For example, "Kill Confirmed" keeps track of "Confirms", "Denies", and "Objectives." 

These are variables that are otherwise unused when not playing "Kill Confirmed". Overall, na.omit is the wrong function to use because it does not consider the possibility that we don't need every column. We should instead use a function that removes observations on a column by column basis.  

```{r}

countNA <- function(dat){
  numNA <- sum(is.na(dat))
  return(numNA)
}

apply(X = COD_dataset, MARGIN = 2, FUN = countNA)
```

## Task 2 - 2 for Jason, 1 for Thadd, 2 for Sammit

### Overview

Explore 5 more variables of your choice. 

- You should explore at least one categorical and at least one quantitative variable.
- Perform a univariate exploration of each variable. (In other words, do not look at the relationship between pairs of variables. Instead, just explore each variable by itself.)

*For each variable, you should...*

1. Identify the variable, the type of variable (categorical, quantitative, etc.), information about the types of values the variable can take, and the amount of missing data within the variable. 
2. Report some summary statistics and make a visualization.
3. *If a variable is categorical...*
    1. Make a table of possible values and include counts or proportions. 
    2. Then, make an appropriate plot to show the distribution of values.
4. *If a variable is quantitative...*
    1. Include some basic summary stats like the mean, median, standard deviation, min, max, etc. 
    2. Make an appropriate plot to show the distribution of 
values.
5. For each variable, write a few sentences to describe what you have learned about the variable. Incorporate the summary statistics and comment on the amount of missingness into your discussion. Remember, the goal is not to make the plot/calculate the summaries. The goal is to use plots and summaries to gain knowledge from the data.
6. Place each discussion next to the relevant plot/summary statistics rather than writing code for all 5 variables first followed by all discussion. If you the discussions are not near the plots, it makes it difficult for your audience to follow your explanations without excessive scrolling



### Deaths (Variable 1)
We were curious about the quantitative variable “Deaths”, which seems to refer to the aggregate number of deaths in a game. Before studying the distribution of the variable, let's first check the number of NA values: 

```{r}
# Filter: Find the number of NA values in Deaths
COD_dataset %>%
  filter(is.na(Deaths)) %>%
  summarize(Number_of_NA = n())
```

...And the statistics collected from the games recorded contained no NA values for Deaths. 


**Summary Statistics:** 

```{r}
# Summary statistics for Deaths variable
COD_dataset %>%
  select(Deaths) %>%
  summarize(min = min(Deaths),
    max = max(Deaths),
    mean = mean(Deaths),
    median = median(Deaths),
    sd = sd(Deaths))
```


**Visualization:**

```{r}
# Histogram of the variable "Death"'s values to show distribution
ggplot(data = COD_dataset,
  mapping = aes(x = Deaths)) +
  geom_histogram(binwidth = 4, fill = "red", color = "black") +
  labs(y = "Number of Rounds",
  title = "Histogram of Death Counts")
```


**Analysis:**

Let's analyze the distribution: based on "Histogram of Death Counts", we can see a concentration of data roughly at the midpoint between 10 and 20 deaths. This is affirmed by the mean and median values reported by the summary statistic table, which is exactly 15.0 and 15 deaths respectively. 

Regarding concentration, the majority of the data seems to lie between 10 and 22 deaths, with roughly $\frac{64 + 76 + 26}{211} = \frac{166}{211} \approx 78.7\%$ of the data— or roughly 3/4ths of the data— lying in this center point. The standard deviation affirms this, suggesting a normal distribution would contain roughly two-thirds of the data between the range 9.87 - 20.13. 

Overall, this distribution is reminiscent of a unimodal, normal distribution with just a slight rightwards skew with a small fraction ($\sim 5$) rounds exceeding 25 Deaths. 



### XPType (Variable 2)

For the second variable, we considered the categorical variable "XPType", which seems to refer to some sort of "bonus" in COD where players would presumably receive a multiplicative/additive/otherwise bonus to their XP gains. 

Based on the dataset, there are 2 different types of XPType:

```{r}
# Table of types of XPType
COD_dataset %>%
  select(XPType) %>%
  distinct()
```

The statistics collected from the rounds recorded contained no NA values for XPType:

```{r}
# Filter: Find the number of NA values in XPType
COD_dataset %>%
  filter(is.na(XPType)) %>%
  summarize(Number_of_NA = n())
```

**Visualization:**

```{r}
# Aggregate of XPType counts per unique XPType
COD_PW_df <- COD_dataset %>%
  select(XPType) %>%
  group_by(XPType) %>%
  summarize(N = n())

# Bar graph visualization of XPType counts
ggplot(data = COD_PW_df,
       mapping = aes(x = XPType, y = N)) +
  geom_bar(stat = "identity", fill = "lightgreen", 
           linewidth = 0.5, colour = "darkgray") +
  labs(y = "Frequency of XPType",
       title = "Bar graph of COD XPType frequencies")

# Table of XPType frequencies
COD_dataset %>%
  select(XPType) %>%
  group_by(XPType) %>%
  summarize(Count = n(), 
            Percent_of_Total = paste0(round(n()/nrow(COD_dataset)*100,2), "%"))
```

**Analysis:** 

There notably isn't enough variety in the sorts of XP bonuses (XPType) to argue any sort of distribution (except, perhaps, Bernoulli). 

Instead, the relative frequency seems to be roughly 1:3, with a little less than 2/5ths (38.86%, to be exact) of rounds played with `Double XP + 10%` XPType and the remainder (a little more than 3/5ths, 61.14% to be exact) of rounds played with a `10% Boost` XPType.The exact counts are that 129 out of 211 rounds were played with a `10% Boost`, and 82 out of 211 rounds were played wiht `Double XP + 10%`.

Why the relative frequencies are not more even or skewed towards `Double XP + 10%`, in spite of this bonus seeming mathematically better with respect to XP gain, is perhaps due to relative availability— the bonus might not be available for the majority of times when players can actually play. 

### Variable 3 - Thadd


### Damage (Variable 4)


We were curious about the quantitative variable “Damage”, which refers to the total amount of damage dealt to opposing players with weapons. The statistics collected from the data set contained no NA values for Damage.


```{r}
# Filter: Find the number of NA values in Damage
COD_dataset %>%
  filter(is.na(Damage)) %>%
  summarize(Number_of_NA = n())
```


**Summary Statistics:** 

```{r}
# Summary statistics for Damage variable
COD_dataset %>%
  select(Damage) %>%
  summarize(min = min(Damage),
    max = max(Damage),
    mean = mean(Damage),
    median = median(Damage),
    sd = sd(Damage),
    Q1 = quantile(Damage, 0.25),
    Q3 = quantile(Damage, 0.75))
```


**Visualization:**

```{r}
# Histogram of the variable "Damage" values to show distribution
ggplot(data = COD_dataset,
  mapping = aes(x = Damage)) +
  geom_histogram(binwidth = 15, fill = "red", color = "black") +
  labs(y = "Number of Rounds",
  title = "Histogram of Damage Dealt")
```


**Analysis:**

Based on the above histogram, we can see a concentration of data between 250 and 550 damage. This is affirmed by the mean and median values reported by the summary statistic table, which is at 415 and 397 damage respectively. We also see an outlier value at 960. Overall, this distribution is unimodal, has a normal distribution with a slight right skew. 


### GameType (Variable 5)


For the fifth variable, we explored the categorical variable "GameType", which refers to the game mode or a new set of objectives the player is meant to accomplish. Game modes can add new more challenging aspects to the game. For example, in Hardpoint an objective is created and your team must control the point for a certain period of time in order to win or score points. 


Based on the dataset, there are 4 different types of GameType's:

```{r}
# Table of types of GameType
COD_dataset %>%
  select(GameType) %>%
  distinct()
```

The statistics collected from the rounds recorded contained no NA values for XPType:

```{r}
#Find the number of NA values in GameType
COD_dataset %>%
  filter(is.na(GameType)) %>%
  summarize(Number_of_NA = n())
```

**Visualization:**

```{r}
# Aggregate of GameType counts per unique GameType
COD_PW_df <- COD_dataset %>%
  select(GameType) %>%
  group_by(GameType) %>%
  summarize(N = n())

# Bar graph visualization of GameType counts
ggplot(data = COD_PW_df,
       mapping = aes(x = GameType, y = N)) +
  geom_bar(stat = "identity", fill = "lightgreen", 
           linewidth = 0.5, colour = "darkgray") +
  labs(y = "Frequency of GameType",
       title = "Bar graph of COD GameType frequencies")

# Table of GameType frequencies
COD_dataset %>%
  select(GameType) %>%
  group_by(GameType) %>%
  summarize(Count = n(), 
            Percent_of_Total = paste0(round(n()/nrow(COD_dataset)*100,2), "%"))
```

**Analysis:** 

As we can see there isn't much variety in GameType, this could be for a variety of reasons. Some could include it being the default game mode recommended to players. 

Instead, the relative frequency seems to be roughly 1:3, with a little less than 2/5ths (38.86%, to be exact) of rounds played with `Double XP + 10%` XPType and the remainder (a little more than 3/5ths, 61.14% to be exact) of rounds played with a `10% Boost` XPType.The exact counts are that 129 out of 211 rounds were played with a `10% Boost`, and 82 out of 211 rounds were played wiht `Double XP + 10%`.

Why the relative frequencies are not more even or skewed towards `Double XP + 10%`, in spite of this bonus seeming mathematically better with respect to XP gain, is perhaps due to relative availability— the bonus might not be available for the majority of times when players can actually play. 




## Task 3 - Jason

**Creating the binary variable `WinLoss`:**

```{r}
# WinLoss = 1 (Win), 0 (Loss)
COD_dataset <- COD_dataset %>%
  # Creating integers PlayerScore and OtherScore
  separate(Result, 
           into = c("PlayerScore", "OtherScore"), 
           sep = "-") %>%
  mutate(PlayerScore = as.integer(PlayerScore)) %>%
  mutate(OtherScore = as.integer(OtherScore)) %>%
  # Restoring "Result" variable
  mutate(Result = COD_dataset$Result) %>%
  mutate(WinLoss = ifelse(PlayerScore > OtherScore, 
                      ifelse(PlayerScore == OtherScore,
                             "Draw", "Win"), "Loss"))
```

**Visualization:**

```{r}
# Aggregate of game state counts per unique game state
COD_WL_df <- COD_dataset %>%
  select(WinLoss) %>%
  group_by(WinLoss) %>%
  summarize(N = n())

# Adding a "Draw" row if it doesn't exist
if(nrow(COD_WL_df) == 2) {
  COD_WL_df <- rbind(COD_WL_df, c("Draw", 0)) %>%
    mutate(N = as.integer(N))
} 

# Bar graph visualization of WinLoss counts
ggplot(data = COD_WL_df,
       mapping = aes(x = fct_relevel(WinLoss, 
                         c("Win", "Draw", "Loss")),
                     y = N, 
                     fill = WinLoss)) +
  geom_bar(stat = "identity", 
           linewidth = 0.5, 
           color = "darkgray") +
  scale_fill_manual(values = c("Win" = "lightgreen", 
                               "Draw" = "lightsalmon", 
                               "Loss" = "lightcoral")) + 
  scale_y_continuous(breaks = seq(0, 120, by = 20)) + 
  labs(x = "Type of Game State (Win/Draw/Loss)",
       y = "Frequency of Wins vs Losses",
       title = "Bar Graph of COD Game State (Win/Draw/Loss) Frequencies")

# Table of XPType frequencies
COD_dataset %>%
  select(WinLoss) %>%
  group_by(WinLoss) %>%
  summarize(Count = n(), 
            Percent_of_Total = paste0(round(n()/nrow(COD_dataset)*100,2), "%"))
```


## Task 4 - Thadd


Question: y = Result (win or loss), x = Score

- i.e. "Is the score related to the result (win/loss)?" 



How to interpret coefficients 


Step A: Exponentiate: calculate 𝑒^𝛽̂1 = 𝑒−0.0088 ≈ 0.991.

Step B: Interpret: As a person’s age increases by 1 year, we expect the odds of surviving to change by a factor of 0.991.

• Note that the explanation reads “change by” rather than “increase by” or “decrease by”
